---
title: "Stage 04 — Clustering Report"
subtitle: "iSN snRNA-seq Pipeline"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    self_contained: true
    dev: ragg_png
knit: (function(input, ...) rmarkdown::render(input, output_dir = "clustering_output", ...))
params:
  project_root: "/scratch/rmlab/rmlab_shared3/tyron/Rscriptv2/iSN/iSN_claude"
  gene_sets: ""
  seed: 123
  track: "decontx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center", dev = "ragg_png")
knitr::opts_knit$set(root.dir = params$project_root)
options(knitr.graphics.rel_path = FALSE)
```

```{r libraries}
library(ggplot2)
library(SeuratDisk)
library(multtest)
library(presto)
library(dplyr)
library(scales)
library(AUCell)
library(miloR)
library(glmGamPoi)
library(future)
library(Seurat)
library(harmony)
library(tidyr)
library(SingleCellExperiment)
library(patchwork)
library(ggrepel)
library(Matrix)
library(RSpectra)
library(viridis)
library(pheatmap)
library(tictoc)
library(scCustomize)
library(scSHC)

plan("multisession", workers = 4)
options(future.globals.maxSize = 30000 * 1024^2, future.seed = TRUE)
```

```{r load-data}
set.seed(params$seed)
out_dir <- file.path(params$project_root, "scripts/04_Clustering/clustering_output")
track_dir <- file.path(out_dir, params$track)
dir.create(track_dir, recursive = TRUE, showWarnings = FALSE)
nf <- 10000
timepoint_colors <- c(Day13 = "#4E8FCA", Day7 = "#D73027", iPSC = "#F5A623")

seu <- readRDS(file.path(out_dir, paste0("04_seu_clustered_", params$track, ".rds")))

# Read markers CSV
all_markers <- read.csv(file.path(track_dir, "04_all_markers_harmony_res0.2.csv"))
```

## Section 1: Sample Overview

```{r sample-overview}
# Cell counts per sample_group
tbl_group <- as.data.frame(table(seu$sample_group))
colnames(tbl_group) <- c("sample_group", "Cell Count")

# Cell counts per sample (timepoint)
tbl_sample <- as.data.frame(table(seu$sample))
colnames(tbl_sample) <- c("Sample (Timepoint)", "Cell Count")

knitr::kable(tbl_group,
             caption = "Cell counts per sample_group",
             row.names = FALSE)

knitr::kable(tbl_sample,
             caption = "Cell counts per timepoint (Day7 / Day13 / iPSC)",
             row.names = FALSE)
```

## Section 2: UMAP — PCA Reduction

```{r umap-pca, fig.width=16, fig.height=7}
sweep_dir <- file.path(track_dir, paste0("nfeatures_", nf))
dir.create(sweep_dir, recursive = TRUE, showWarnings = FALSE)

centroids_pca_cl <- as.data.frame(Embeddings(seu, "umap.pca"))
centroids_pca_cl$pca_res.0.2 <- seu$pca_res.0.2
centroids_pca_cl <- aggregate(. ~ pca_res.0.2, centroids_pca_cl, mean)

p_pca_clusters <- DimPlot(seu,
                           reduction = "umap.pca",
                           group.by  = "pca_res.0.2",
                           label     = FALSE,
                           pt.size   = 0.5) +
  ggtitle(paste0("iSN PCA clusters (res 0.2, nf=", nf, ")")) +
  xlab("UMAP 1") + ylab("UMAP 2") +
  theme(
    plot.title   = element_text(size = 15, face = "bold", hjust = 0.5),
    axis.title   = element_text(size = 13, face = "bold"),
    axis.text    = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 12)
  ) +
  scale_x_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  scale_y_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  geom_text_repel(
    data = centroids_pca_cl,
    aes(x = umappca_1, y = umappca_2, label = pca_res.0.2),
    size = 5, color = "black", segment.color = "grey50",
    box.padding = 0.8, max.overlaps = Inf
  )

centroids_pca_samp <- as.data.frame(Embeddings(seu, "umap.pca"))
centroids_pca_samp$sample <- seu$sample
centroids_pca_samp <- aggregate(. ~ sample, centroids_pca_samp, mean)

p_pca_samples <- DimPlot(seu,
                          reduction = "umap.pca",
                          group.by  = "sample",
                          label     = FALSE,
                          pt.size   = 0.5) +
  ggtitle(paste0("iSN PCA — sample (nf=", nf, ")")) +
  xlab("UMAP 1") + ylab("UMAP 2") +
  theme(
    plot.title   = element_text(size = 15, face = "bold", hjust = 0.5),
    axis.title   = element_text(size = 13, face = "bold"),
    axis.text    = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 12)
  ) +
  scale_color_manual(values = timepoint_colors) +
  scale_x_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  scale_y_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  geom_text_repel(
    data = centroids_pca_samp,
    aes(x = umappca_1, y = umappca_2, label = sample),
    size = 5, color = "black", segment.color = "grey50",
    box.padding = 0.8, max.overlaps = Inf
  )

patchwork_pca <- p_pca_clusters | p_pca_samples
ggsave(file.path(sweep_dir, paste0("dimplot_pca_nf", nf, ".pdf")),
       plot = patchwork_pca, width = 16, height = 7)
patchwork_pca
```

## Section 3: UMAP — Harmony Reduction

```{r umap-harmony, fig.width=16, fig.height=7}
centroids_harm_cl <- as.data.frame(Embeddings(seu, "umap.harmony"))
centroids_harm_cl$harmony_res.0.2 <- seu$harmony_res.0.2
centroids_harm_cl <- aggregate(. ~ harmony_res.0.2, centroids_harm_cl, mean)

p_harmony_clusters <- DimPlot(seu,
                               reduction = "umap.harmony",
                               group.by  = "harmony_res.0.2",
                               label     = FALSE,
                               pt.size   = 0.5) +
  ggtitle(paste0("iSN Harmony clusters (res 0.2, nf=", nf, ")")) +
  xlab("UMAP 1") + ylab("UMAP 2") +
  theme(
    plot.title   = element_text(size = 15, face = "bold", hjust = 0.5),
    axis.title   = element_text(size = 13, face = "bold"),
    axis.text    = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 12)
  ) +
  scale_x_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  scale_y_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  geom_text_repel(
    data = centroids_harm_cl,
    aes(x = umapharmony_1, y = umapharmony_2, label = harmony_res.0.2),
    size = 5, color = "black", segment.color = "grey50",
    box.padding = 0.8, max.overlaps = Inf
  )

centroids_harm_samp <- as.data.frame(Embeddings(seu, "umap.harmony"))
centroids_harm_samp$sample <- seu$sample
centroids_harm_samp <- aggregate(. ~ sample, centroids_harm_samp, mean)

p_harmony_samples <- DimPlot(seu,
                              reduction = "umap.harmony",
                              group.by  = "sample",
                              label     = FALSE,
                              pt.size   = 0.5) +
  ggtitle(paste0("iSN Harmony — sample (nf=", nf, ")")) +
  xlab("UMAP 1") + ylab("UMAP 2") +
  theme(
    plot.title   = element_text(size = 15, face = "bold", hjust = 0.5),
    axis.title   = element_text(size = 13, face = "bold"),
    axis.text    = element_text(size = 13),
    legend.title = element_text(size = 12),
    legend.text  = element_text(size = 12)
  ) +
  scale_color_manual(values = timepoint_colors) +
  scale_x_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  scale_y_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
  geom_text_repel(
    data = centroids_harm_samp,
    aes(x = umapharmony_1, y = umapharmony_2, label = sample),
    size = 5, color = "black", segment.color = "grey50",
    box.padding = 0.8, max.overlaps = Inf
  )

patchwork_harmony <- p_harmony_clusters | p_harmony_samples
ggsave(file.path(sweep_dir, paste0("dimplot_harmony_nf", nf, ".pdf")),
       plot = patchwork_harmony, width = 16, height = 7)
patchwork_harmony
```

## Section 4: scSHC vs Harmony Cluster Comparison

```{r scshc-comparison, eval=FALSE, fig.width=16, fig.height=7}
# scSHC disabled — OOM on 65k cells; scshc_clusters column not present
# p_scshc <- DimPlot_scCustom(seu,
#                              reduction = "umap.harmony",
#                              group.by  = "scshc_clusters",
#                              label     = TRUE,
#                              pt.size   = 0.5) +
#   ggtitle("scSHC clusters") +
#   xlab("UMAP 1") + ylab("UMAP 2") +
#   theme(
#     plot.title   = element_text(size = 15, face = "bold", hjust = 0.5),
#     axis.title   = element_text(size = 13, face = "bold"),
#     axis.text    = element_text(size = 13),
#     legend.title = element_text(size = 12),
#     legend.text  = element_text(size = 12)
#   ) +
#   scale_x_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
#   scale_y_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4))
#
# p_seurat_res <- DimPlot_scCustom(seu,
#                                   reduction = "umap.harmony",
#                                   group.by  = "harmony_res.0.2",
#                                   label     = TRUE,
#                                   pt.size   = 0.5) +
#   ggtitle("Seurat Harmony clusters (res 0.2)") +
#   xlab("UMAP 1") + ylab("UMAP 2") +
#   theme(
#     plot.title   = element_text(size = 15, face = "bold", hjust = 0.5),
#     axis.title   = element_text(size = 13, face = "bold"),
#     axis.text    = element_text(size = 13),
#     legend.title = element_text(size = 12),
#     legend.text  = element_text(size = 12)
#   ) +
#   scale_x_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
#   scale_y_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4))
#
# p_compare <- p_scshc | p_seurat_res
# ggsave(file.path(track_dir, "scshc_vs_harmony_res0.2.pdf"),
#        plot = p_compare, width = 16, height = 7)
# p_compare
```

## Section 5: Dimensionality Reduction Diagnostics

```{r elbow-jackstraw, fig.width=12, fig.height=5}
p_elbow_final <- ElbowPlot(seu, ndims = 80) + theme_bw()
# p_jackstraw   <- JackStrawPlot(seu, dims = 1:80)  # JackStraw not run — disabled

p_elbow_final
# p_jackstraw
```

## Section 6: AUCell — Cell Cycle and G2M Gene Sets

```{r aucell-run}
cellcycle_genes <- c(
  "TOP2A",
  "MCM2",
  "MCM3",
  "MCM5",
  "MCM6",
  "MKI67"
)

g2m.genes <- c(
  "ATF5",
  "AURKA",
  "AURKB",
  "BARD1",
  "BIRC5",
  "BRCA2",
  "BUB1",
  "CCNA2",
  "CCNB2",
  "CCND1",
  "CDC20",
  "CDC45",
  "CDC6",
  "CDK1",
  "CDK2",
  "CDK4",
  "CDKN3",
  "CENPA",
  "CENPE",
  "CENPF",
  "CHAF1A",
  "CHEK1",
  "CKS1B",
  "CKS2",
  "E2F1",
  "E2F2",
  "EGF",
  "ESPL1",
  "EXO1",
  "FBXO5",
  "GINS2",
  "HMGN2",
  "HMMR",
  "KIF11",
  "KIF15",
  "KIF20B",
  "KIF23",
  "KIF2C",
  "KIF4A",
  "KNL1",
  "LMNB1",
  "MAD2L1",
  "MAP3K20",
  "MCM2",
  "MCM3",
  "MCM4",
  "MCM5",
  "MCM6",
  "MKI67",
  "MT2A",
  "MYBL2",
  "NDC80",
  "NEK2",
  "NOTCH2",
  "NUSAP1",
  "ORC6",
  "PBK",
  "PCNA",
  "PLK1",
  "PLK4",
  "POLA2",
  "POLQ",
  "PTTG1",
  "RAD54L",
  "SMC4",
  "STIL",
  "TACC3",
  "TOP2A",
  "TPX2",
  "TROAP",
  "TTK",
  "UBE2C"
)

DefaultLayer(seu[["RNA"]]) <- "data"
exprMatrix <- as.matrix(seu[["RNA"]]$data)

cells_rankings <- AUCell_buildRankings(exprMatrix, plotStats = FALSE)

cells_AUC_cellcycle <- AUCell_calcAUC(cellcycle_genes, cells_rankings, aucMaxRank = 20000)
cells_AUC_g2m       <- AUCell_calcAUC(g2m.genes,       cells_rankings, aucMaxRank = 20000)

dir.create(file.path(track_dir, "aucell"), recursive = TRUE, showWarnings = FALSE)

pdf(file.path(track_dir, "aucell", "aucell_thresholds_cellcycle.pdf"))
cells_assignment_cellcycle <- AUCell_exploreThresholds(cells_AUC_cellcycle,
                                                       plotHist = TRUE,
                                                       assign   = TRUE)
dev.off()

pdf(file.path(track_dir, "aucell", "aucell_thresholds_g2m.pdf"))
cells_assignment_g2m <- AUCell_exploreThresholds(cells_AUC_g2m,
                                                 plotHist = TRUE,
                                                 assign   = TRUE)
dev.off()
```

_AUCell threshold histograms saved to `aucell/aucell_thresholds_cellcycle.pdf` and `aucell/aucell_thresholds_g2m.pdf`_

```{r aucell-helpers}
# ggplot2 AUCell scatter plot
plot_aucell_simple <- function(seurat_obj, cells_AUC, cells_assignment,
                               gene_set_name, reduction = "umap.harmony",
                               only_passing = FALSE, only_failing = FALSE) {
  emb <- as.data.frame(seurat_obj@reductions[[reduction]]@cell.embeddings)
  umap_cols <- colnames(emb)
  selectedThreshold <- unname(cells_assignment[[1]]$aucThr$selected)
  auc_scores <- getAUC(cells_AUC)[1, ]
  passThreshold <- auc_scores > selectedThreshold
  n_passed <- sum(passThreshold)
  emb$status <- ifelse(passThreshold[rownames(emb)], "Above threshold", "Below threshold")

  if (only_passing) {
    emb <- emb[emb$status == "Above threshold", , drop = FALSE]
    subtitle <- paste0(n_passed, " cells above threshold")
  } else if (only_failing) {
    emb <- emb[emb$status == "Below threshold", , drop = FALSE]
    subtitle <- paste0(ncol(seurat_obj) - n_passed, " cells below threshold")
  } else {
    subtitle <- paste0(n_passed, " cells above threshold (pink)")
  }

  ggplot(emb, aes(x = .data[[umap_cols[1]]], y = .data[[umap_cols[2]]], color = status)) +
    geom_point(size = 0.5, alpha = 0.7) +
    scale_color_manual(values = c("Above threshold" = "deeppink", "Below threshold" = "gray80"),
                       drop = FALSE) +
    scale_x_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
    scale_y_continuous(limits = c(-16, 16), breaks = seq(-16, 16, 4)) +
    labs(title = gene_set_name, subtitle = subtitle,
         x = "UMAP_1", y = "UMAP_2", color = "AUC Status") +
    theme_classic() +
    theme(plot.title = element_text(face = "bold", hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5),
          legend.position = "right")
}

# Bar chart of cells above/below AUC threshold per cluster
plot_aucell_cluster_counts <- function(seurat_obj, cells_AUC, cells_assignment,
                                       gene_set_name,
                                       cluster_col = "harmony_res.0.2") {

  selectedThreshold <- unname(cells_assignment[[1]]$aucThr$selected)
  auc_scores        <- getAUC(cells_AUC)[1, ]
  passThreshold     <- auc_scores > selectedThreshold

  df <- data.frame(
    cell    = names(auc_scores),
    cluster = seurat_obj@meta.data[[cluster_col]][
      match(names(auc_scores), rownames(seurat_obj@meta.data))
    ],
    status  = ifelse(passThreshold, "Above threshold", "Below threshold"),
    stringsAsFactors = FALSE
  ) %>%
    filter(!is.na(cluster)) %>%
    mutate(cluster = factor(cluster))

  counts_df <- df %>%
    group_by(cluster, status) %>%
    summarise(n = n(), .groups = "drop")

  fisher_results <- df %>%
    group_by(cluster) %>%
    do({
      current_cluster <- unique(.$cluster)
      a <- sum(.$status == "Above threshold")
      b <- sum(.$status == "Below threshold")
      others <- df %>% filter(cluster != current_cluster)
      c <- sum(others$status == "Above threshold")
      d <- sum(others$status == "Below threshold")
      mat  <- matrix(c(a, b, c, d), nrow = 2)
      pval <- fisher.test(mat)$p.value
      data.frame(p = pval)
    }) %>%
    ungroup()

  fisher_results$p_adj <- p.adjust(fisher_results$p, method = "BH")
  fisher_results$stars <- cut(
    fisher_results$p_adj,
    breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
    labels = c("***", "**", "*", "ns")
  )

  max_counts <- counts_df %>%
    group_by(cluster) %>%
    summarise(max_n = max(n), .groups = "drop")

  fisher_results <- left_join(fisher_results, max_counts, by = "cluster")

  global_max             <- max(counts_df$n)
  counts_df$y_label      <- counts_df$n + global_max * 0.015
  fisher_results$y_pos   <- fisher_results$max_n + global_max * 0.12

  dodge <- position_dodge(width = 1)

  p <- ggplot(counts_df, aes(x = cluster, y = n, fill = status)) +
    geom_bar(stat = "identity", position = dodge, width = 0.8, alpha = 0.85) +
    geom_text(aes(y = y_label, label = n),
              position = dodge, size = 3, fontface = "bold") +
    geom_text(data = fisher_results %>% filter(stars != "ns"),
              aes(x = cluster, y = y_pos, label = stars),
              inherit.aes = FALSE, size = 5, fontface = "bold") +
    geom_text(data = fisher_results %>% filter(stars == "ns"),
              aes(x = cluster, y = y_pos, label = stars),
              inherit.aes = FALSE, size = 3.5, color = "gray40") +
    scale_fill_manual(values = c("Above threshold" = "deeppink",
                                 "Below threshold"  = "gray70")) +
    scale_y_continuous(limits = c(0, global_max * 1.4),
                       expand = expansion(mult = c(0, 0))) +
    labs(title = paste0(gene_set_name,
                        " — Cells Above/Below AUC Threshold per Cluster"),
         x    = "Cluster", y = "Number of Cells", fill = "AUC Status") +
    theme_classic() +
    theme(axis.text.x     = element_text(angle = 45, hjust = 1),
          legend.position  = "top",
          plot.title       = element_text(face = "bold", hjust = 0.5))

  return(p)
}

# Bar chart by timepoint
make_aucell_timepoint_barchart <- function(cells_AUC, cells_assignment,
                                            seurat_obj, timepoints, tp_colors,
                                            title_label) {

  auc_scores        <- getAUC(cells_AUC)[1, ]
  selectedThreshold <- unname(cells_assignment[[1]]$aucThr$selected)

  df_all <- data.frame(
    cell      = names(auc_scores),
    timepoint = seurat_obj@meta.data[
      match(names(auc_scores), rownames(seurat_obj@meta.data)), "sample"
    ],
    status = ifelse(auc_scores > selectedThreshold, "Above threshold", "Below threshold"),
    stringsAsFactors = FALSE
  ) %>%
    filter(timepoint %in% timepoints) %>%
    mutate(timepoint = factor(timepoint, levels = timepoints))

  counts_df <- df_all %>%
    group_by(timepoint, status) %>%
    summarise(n = n(), .groups = "drop")

  fisher_results <- lapply(timepoints, function(tp) {
    a <- sum(df_all$timepoint == tp & df_all$status == "Above threshold")
    b <- sum(df_all$timepoint == tp & df_all$status == "Below threshold")
    others <- df_all %>% filter(timepoint != tp)
    c <- sum(others$status == "Above threshold")
    d <- sum(others$status == "Below threshold")
    mat  <- matrix(c(a, b, c, d), nrow = 2)
    pval <- fisher.test(mat)$p.value
    data.frame(timepoint = tp, p = pval)
  }) %>% bind_rows()

  fisher_results$p_adj <- p.adjust(fisher_results$p, method = "BH")
  fisher_results$stars <- cut(
    fisher_results$p_adj,
    breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
    labels = c("***", "**", "*", "ns")
  )

  global_max <- max(counts_df$n)

  max_per_tp <- counts_df %>%
    group_by(timepoint) %>%
    summarise(max_n = max(n), .groups = "drop")

  fisher_results <- left_join(fisher_results, max_per_tp, by = "timepoint")
  fisher_results$y_pos <- fisher_results$max_n + global_max * 0.08
  counts_df$y_label    <- counts_df$n + global_max * 0.01

  dodge <- position_dodge(width = 0.8)

  p <- ggplot(counts_df, aes(x = timepoint, y = n, fill = status)) +
    geom_bar(stat = "identity", position = dodge, width = 0.7, alpha = 0.85) +
    geom_text(aes(y = y_label, label = n),
              position = dodge, size = 3, fontface = "bold") +
    geom_text(data = fisher_results %>% filter(stars != "ns"),
              aes(x = timepoint, y = y_pos, label = stars),
              inherit.aes = FALSE, size = 5, fontface = "bold") +
    geom_text(data = fisher_results %>% filter(stars == "ns"),
              aes(x = timepoint, y = y_pos, label = stars),
              inherit.aes = FALSE, size = 3.5, color = "gray40") +
    scale_fill_manual(values = c("Above threshold" = "deeppink",
                                 "Below threshold"  = "gray70")) +
    scale_y_continuous(limits = c(0, global_max * 1.3),
                       expand = expansion(mult = c(0, 0))) +
    labs(title = title_label, x = "Timepoint",
         y = "Number of Cells", fill = "AUC Status") +
    theme_classic() +
    theme(plot.title       = element_text(face = "bold", hjust = 0.5),
          legend.position  = "top",
          axis.text.x      = element_text(size = 12, face = "bold",
                                          color = tp_colors[levels(counts_df$timepoint)]))

  return(p)
}
```

### AUCell: Cell Cycle — UMAP scatter plots

```{r aucell-cellcycle-umap, fig.width=10, fig.height=8}
plot_aucell_simple(seu, cells_AUC_cellcycle, cells_assignment_cellcycle,
                   "CellCycle", reduction = "umap.harmony")

plot_aucell_simple(seu, cells_AUC_cellcycle, cells_assignment_cellcycle,
                   "CellCycle", reduction = "umap.harmony", only_passing = TRUE)

plot_aucell_simple(seu, cells_AUC_cellcycle, cells_assignment_cellcycle,
                   "CellCycle", reduction = "umap.harmony", only_failing  = TRUE)
```

### AUCell: G2M — UMAP scatter plots

```{r aucell-g2m-umap, fig.width=10, fig.height=8}
plot_aucell_simple(seu, cells_AUC_g2m, cells_assignment_g2m,
                   "G2M", reduction = "umap.harmony")

plot_aucell_simple(seu, cells_AUC_g2m, cells_assignment_g2m,
                   "G2M", reduction = "umap.harmony", only_passing = TRUE)

plot_aucell_simple(seu, cells_AUC_g2m, cells_assignment_g2m,
                   "G2M", reduction = "umap.harmony", only_failing  = TRUE)
```

### AUCell: Cluster bar charts

```{r aucell-cluster-barcharts, fig.width=10, fig.height=6}
timepoints <- c("Day13", "Day7", "iPSC")
tp_colors  <- timepoint_colors

p_cc_clusters <- plot_aucell_cluster_counts(seu, cells_AUC_cellcycle,
                                             cells_assignment_cellcycle,
                                             gene_set_name = "CellCycle",
                                             cluster_col   = "harmony_res.0.2")
ggsave(file.path(track_dir, "aucell", "aucell_cellcycle_cluster_barchart.pdf"),
       plot = p_cc_clusters, width = 10, height = 6)

p_g2m_clusters <- plot_aucell_cluster_counts(seu, cells_AUC_g2m,
                                              cells_assignment_g2m,
                                              gene_set_name = "G2M",
                                              cluster_col   = "harmony_res.0.2")
ggsave(file.path(track_dir, "aucell", "aucell_g2m_cluster_barchart.pdf"),
       plot = p_g2m_clusters, width = 10, height = 6)

p_cc_clusters
p_g2m_clusters
```

### AUCell: Timepoint bar charts

```{r aucell-timepoint-barcharts, fig.width=8, fig.height=6}
p_cc_tp <- make_aucell_timepoint_barchart(
  cells_AUC_cellcycle, cells_assignment_cellcycle, seu, timepoints, tp_colors,
  "Cell Cycle Gene Set — Cells Above/Below AUC Threshold by Timepoint"
)
ggsave(file.path(track_dir, "aucell", "aucell_cellcycle_sample_barchart.pdf"),
       plot = p_cc_tp, width = 8, height = 6)

p_g2m_tp <- make_aucell_timepoint_barchart(
  cells_AUC_g2m, cells_assignment_g2m, seu, timepoints, tp_colors,
  "G2M Gene Set — Cells Above/Below AUC Threshold by Timepoint"
)
ggsave(file.path(track_dir, "aucell", "aucell_g2m_sample_barchart.pdf"),
       plot = p_g2m_tp, width = 8, height = 6)

p_cc_tp
p_g2m_tp
```

## Section 7: Marker Gene Expression

```{r marker-featureplot, fig.width=16, fig.height=12}
isN_markers <- c(
  "TUBB3", "PRPH", "SNAP25",
  "CALCA", "TRPV1",
  "MRGPRD",
  "NTRK2", "NTRK3",
  "POU5F1", "SOX2", "NANOG"
)

p_feature_harmony <- FeaturePlot(seu, features = isN_markers,
                                  reduction = "umap.harmony", ncol = 4) &
  scale_color_viridis_c(option = "plasma")
ggsave(file.path(track_dir, "featureplot_markers_harmony.pdf"),
       plot = p_feature_harmony, width = 16, height = 12)
p_feature_harmony
```

```{r marker-dotplot, fig.width=10, fig.height=6}
p_dot <- DotPlot(seu, features = isN_markers, group.by = "harmony_res.0.2") +
  RotatedAxis() +
  scale_color_gradient(low = "white", high = "red",
                       limits = c(-1, 3), oob = squish)
ggsave(file.path(track_dir, "dotplot_isN_markers.pdf"),
       plot = p_dot, width = 10, height = 6)
p_dot
```

## Section 8: Module Scores (Hardcoded Gene Sets)

```{r module-scores, fig.width=14, fig.height=9}
score_features <- c("pan_neuronal_score1", "peptidergic_score1",
                    "non_peptidergic_score1", "trkbc_score1", "ipsc_score1")

p_scores <- FeaturePlot(seu, features = score_features,
                        reduction = "umap.harmony", ncol = 3,
                        min.cutoff = "q05", max.cutoff = "q95") &
  scale_color_viridis_c(option = "plasma")
ggsave(file.path(track_dir, "featureplot_module_scores.pdf"),
       plot = p_scores, width = 14, height = 9)
p_scores
```

### Section 8.1: User-Specified Gene Sets

```{r user-gene-sets}
args_gene_sets <- params$gene_sets
if (!is.null(args_gene_sets) && nchar(args_gene_sets) > 0) {
  sets_raw <- strsplit(args_gene_sets, ";")[[1]]
  gene_set_list <- lapply(sets_raw, function(s) {
    parts <- strsplit(s, "=")[[1]]
    list(name = parts[1], genes = strsplit(parts[2], ",")[[1]])
  })
  for (gs in gene_set_list) {
    set_name  <- gs$name
    set_genes <- gs$genes
    score_col <- paste0("score_", set_name)
    seu_scored <- tryCatch(
      AddModuleScore(seu, features = list(set_genes), name = score_col),
      error = function(e) {
        message("Skipping gene set '", set_name, "': ", conditionMessage(e))
        NULL
      }
    )
    if (is.null(seu_scored)) next
    seu <- seu_scored
    seu[[score_col]] <- seu[[paste0(score_col, "1")]]
    seu[[paste0(score_col, "1")]] <- NULL
    p_umap <- FeaturePlot(seu, features = score_col,
                          reduction = "umap.harmony", pt.size = 0.3) +
              scale_colour_viridis_c() +
              ggtitle(paste("Module score:", set_name))
    print(p_umap)
    p_vln <- VlnPlot(seu, features = score_col,
                     group.by = "harmony_res.0.2", pt.size = 0) +
             ggtitle(paste("Module score:", set_name)) + NoLegend()
    print(p_vln)
  }
}
```

## Section 9: Violin Plots (iSN Marker Genes)

```{r violin-plots, fig.width=8, fig.height=5}
iqr_summary <- function(x) {
  data.frame(
    y    = median(x),
    ymin = quantile(x, 0.25),
    ymax = quantile(x, 0.75)
  )
}

violin_genes <- c(
  "TUBB3", "PRPH", "SNAP25", "CALCA", "TRPV1", "MRGPRD",
  "NTRK2", "NTRK3", "POU5F1", "SOX2", "NANOG",
  "MKI67", "TOP2A", "MCM2"
)

Idents(seu) <- seu$harmony_res.0.2

violin_genes_present <- intersect(violin_genes, rownames(seu[["RNA"]]))
for (gene in violin_genes_present) {
  df_vln <- FetchData(seu, vars = c(gene, "harmony_res.0.2"))
  colnames(df_vln)[1] <- "expr"

  p_vln <- ggplot(df_vln, aes(x = harmony_res.0.2, y = expr,
                               fill = harmony_res.0.2)) +
    geom_violin(scale = "width", trim = TRUE) +
    stat_summary(fun.data = iqr_summary, geom = "crossbar",
                 width = 0.3, color = "black") +
    scale_fill_brewer(palette = "Set3") +
    theme_classic() +
    theme(legend.position = "none",
          plot.title      = element_text(face = "bold", hjust = 0.5),
          axis.text.x     = element_text(angle = 45, hjust = 1)) +
    labs(title = gene, x = "Cluster (harmony_res.0.2)", y = "Expression")

  ggsave(file.path(track_dir, paste0("violin_", gene, ".pdf")),
         plot = p_vln, width = 8, height = 5)
  print(p_vln)
}
```

## Section 10: Sample Composition

```{r pie-chart, fig.width=7, fig.height=7}
group_df <- seu@meta.data %>%
  group_by(sample) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(percentage = count / sum(count) * 100)

group_df$sample <- factor(group_df$sample, levels = c("iPSC", "Day7", "Day13"))

group_df <- group_df %>%
  arrange(sample) %>%
  mutate(
    label = paste0(round(percentage, 1), "%\n(n=", count, ")"),
    csum  = rev(cumsum(rev(percentage))),
    ypos  = percentage / 2 + lead(csum, default = 0)
  )

p_pie <- ggplot(group_df, aes(x = "", y = percentage, fill = sample)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  theme_void() +
  geom_text(aes(y = ypos, label = label),
            fontface = "bold", size = 4, color = "black") +
  scale_fill_manual(values = c("iPSC"  = "#F5A623",
                               "Day7"  = "#D73027",
                               "Day13" = "#4E8FCA")) +
  ggtitle("iSN samples — Day7 / Day13 / iPSC") +
  theme(plot.title      = element_text(hjust = 0.5, size = 18, face = "bold"),
        legend.title    = element_blank(),
        legend.text     = element_text(size = 12))

ggsave(file.path(track_dir, "piechart_by_sample.pdf"),
       plot = p_pie, width = 7, height = 7)
p_pie
```

## Section 11: Top Marker Gene Heatmap

```{r heatmap, fig.width=14, fig.height=10}
Idents(seu) <- seu$harmony_res.0.2

top5 <- all_markers %>%
  group_by(cluster) %>%
  slice_max(order_by = avg_log2FC, n = 5) %>%
  ungroup()

DefaultLayer(seu[["RNA"]]) <- "data"

set.seed(params$seed)
cells.use <- unlist(
  lapply(split(Cells(seu), Idents(seu)),
         function(x) sample(x, min(length(x), 100)))
)

p_heatmap <- DoHeatmap(seu,
                       features  = unique(top5$gene),
                       cells     = cells.use,
                       group.by  = "harmony_res.0.2",
                       raster    = TRUE) +
  ggtitle("Top 5 Marker Genes per Cluster (Downsampled Heatmap)")

ggsave(file.path(track_dir, "04_heatmap_top5_markers.pdf"),
       plot = p_heatmap, width = 14, height = 10)
p_heatmap
```

## Session Info

```{r session-info}
sessionInfo()
```
